### 2014-09-25 ApTranscriptome_TR.Rmd
updated TR

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods) # required by `knitcitations` when file is `knit` from command-line
library(reshape2)
library(RColorBrewer)

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
#biocLite("GOSemSim")
library(topGO) 
library(Rgraphviz)	
library(GOSemSim)

@@

```{r responsive_transcripts}
interaction.lms <- RxNlmAIC[which(Map(grepFunc, RxNlmAIC, term = "colonyAr:") == TRUE)]
other.lms <- RxNlmAIC[setdiff(names(RxNlmAIC), names(interaction.lms))]
temperature.lms <- other.lms[which(Map(grepFunc, other.lms, term = "val") == TRUE)]
colony.lms <- other.lms[setdiff(names(other.lms), names(temperature.lms))]
responsive.lms <- c(temperature.lms, interaction.lms)
rm(other.lms)
```

@@

```{r mh_test}
mh.test <- mh_test(tt2)

@@

# proportion of test statistic due to off-diagonals
dsum <- sum(abs(d1), abs(d2), abs(d3), abs(d4), abs(d5))
d1/dsum
d2/dsum
d3/dsum
d4/dsum
d5/dsum

@@

qplot(x=Apic, y=Acar, data=md, fill=value, geom="tile", ylim = rev(levels(md$Acar))) + 
  scale_fill_gradient2(limits=c(-600, 600)) 
```

@@

## Species comparison

@@

Of the responsive transcripts, `r prettyNum(length(which(responsive.lms.ann.type$best.hit.to.nr == "-"))/length(responsive.lms.ann.type$best.hit.to.nr)*100)`% are annotated, while `r prettyNum(length(which(annotation.table$best.hit.to.nr == "-"))/nrow(annotation.table) * 100)`% of all transcripts are annotated.

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, hsp_responsive, hsp_all, Ap.BP.resultParentChild, resp.stress.responsive.count, T_on.high.ttest, T_on.low.ttest, T_off.int.high.ttest, T_off.int.low.ttest, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, Ar.int.A22.lowbim.gsea, file = "doc/ms.RData")
```

>>>>

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods) # required by `knitcitations` when file is `knit` from command-line
library(reshape2)
library(RColorBrewer)
library(coin)

# Load Bioconductor libraries
source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
#biocLite("GOSemSim")
library(topGO) 
library(Rgraphviz)	
library(GOSemSim)

@@

```{r responsive_transcripts}
interaction.lms <- RxNlmAIC[which(Map(grepFunc, RxNlmAIC, term = "colonyAr:") == TRUE)]
other.lms <- RxNlmAIC[setdiff(names(RxNlmAIC), names(interaction.lms))]
temperature.lms <- other.lms[which(Map(grepFunc, other.lms, term = "val") == TRUE)]
colony.lms <- other.lms[setdiff(names(other.lms), names(temperature.lms))]
responsive.lms <- c(temperature.lms, interaction.lms)
rm(other.lms)

quadratic.lms <- RxNlmAIC[which(Map(grepFunc, RxNlmAIC, term = "2") == TRUE)]
```

@@

```{r mh_test}
mh.test <- mh_test(as.table(tt2))

@@

# proportion of test statistic due to off-diagonals
dsum <- sum(abs(d1), abs(d2), abs(d3), abs(d4), abs(d5))
abs(d1)/dsum
abs(d2)/dsum
abs(d3)/dsum
abs(d4)/dsum
abs(d5)/dsum

@@

mh_plot <- qplot(x=Apic, y=Acar, data=md, fill=value, geom="tile", ylim = rev(levels(md$Acar))) + 
  scale_fill_gradient2(limits=c(-600, 600)) 
mh_plot
```

@@

An interesting observation from the matched observations plot (Fig. ...) is that for **NotResp** genes in *A. carolinensis* the **High** and **Bimodal** categories in *A. picea* are over-represented. Finding that the mean expression level of these genes in *A. carolinensis* is greater than the expression level at the optimum temperature (19.5°C) in *A. picea* would be consistent with genetic assimilation.

```{r gen_assimilation_high}
Ap.high.Ac.nr <- Ap.response.type[which(Ap.response.type$A22.type == "High" & Ap.response.type$Ar.type == "NotResp"), ]

# get mean expression level for Ar transcripts
Ap.high.Ac.nr.TPM <- TPM.dt.sub[Ap.high.Ac.nr$Transcript]
Ap.high.Ac.nr.mean.exp.NR <- ddply(Ap.high.Ac.nr.TPM, .(Transcript), summarize, meanTPM = mean(TPM))

# get expression at optimum for A22 transcripts

Ap.high.Ac.nr.opt.exp <- Ap.high.Ac.nr$A22.opt

t.test(Ap.high.Ac.nr.mean.exp.NR$meanTPM, Ap.high.Ac.nr.opt.exp)
```

Contrary to expectations, mean expression of the non-responsive *A. carolinensis* genes is less than expression of the **High** genes in *A. picea*.

Next, I check the same for the **Bimodal** genes in *A. picea*.

```{r gen_assimilation_bim}
Ap.bim.Ac.nr <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "NotResp"), ]

# get mean expression level for Ar transcripts
Ap.bim.Ac.nr.TPM <- TPM.dt.sub[Ap.bim.Ac.nr$Transcript]
Ap.bim.Ac.nr.mean.exp.NR <- ddply(Ap.bim.Ac.nr.TPM, .(Transcript), summarize, meanTPM = mean(TPM))

# get expression at optimum for A22 transcripts
Ap.bim.Ac.nr.opt.exp <- Ap.bim.Ac.nr$A22.opt

t.test(Ap.bim.Ac.nr.mean.exp.NR$meanTPM, Ap.bim.Ac.nr.opt.exp)
```

As with the **High** transcripts, expression much higher for the responsive *A. picea* transcripts than for the non-responsive *A. carolinensis* transcripts.

## Species comparisons

@@

Of the responsive transcripts, `r round(length(which(responsive.lms.ann.type$best.hit.to.nr != "-"))/length(responsive.lms.ann.type$best.hit.to.nr)*100, 0)`% are annotated, while `r round(length(which(annotation.table$best.hit.to.nr != "-"))/nrow(annotation.table) * 100, 0)`% of all transcripts are annotated.

@@

```{r rData, echo=FALSE}
# save RData for manuscript file
save(TPM.dt.sub, RxNpval, sigtable, chi1, resp.TPM.dt.sub, Ap.type.table, tt2, tt3, GO.stress.Xsq, GO0006950.responsive, Ap.df, w1, w2, w3, Ap.T_on, hsp_responsive, hsp_all, mh_plot, Ap.BP.resultParentChild, resp.stress.responsive.count, T_on.high.ttest, T_on.low.ttest, T_off.int.high.ttest, T_off.int.low.ttest, Ap.gsea.intersect, Ar.int.A22.bim.gsea, Ar.int.A22.low.gsea, Ar.int.A22.lowbim.gsea, quadratic.lms, file = paste("doc/ms_", Sys.Date(), ".RData", sep = ""))
```
